Market-based Mechanisms for Resource Scheduling

### 2013-12-18 paper/paper.tex
Added property table, added eval graphs

\usepackage{mathtools}
\usepackage{graphicx}

@@

In section \ref{sec:design}, we describe our protocol and the most important design decisions
we made. In particular, we describe the mechanisms we used to ensure that our
protocol remains backwards-compatible, and can be applied without any change to
the clients’ code.

@@

This example clearly shows that in this case, the mechanism provided by Mesos is
\emph{not} strategy-proof. Worse, it’s not Pareto efficient if people are
honest, and need not be Pareto efficient if clients behave optimally. Similar
things happen in the other examples we mentioned, or in most of the similar
alternatives which we can envision:
\begin{enumerate}
  \item When one client has a constant stream of work and another has varying
    traffic, Mesos encourages both to constantly use up their allocated share of
    the network (unless they have nothing to do with the space). But an
    efficient allocation would give the client with varying workload more
    resources when it has more work. 
  \item Mesos gives a valuable resource to whatever client happens to be using
    the fewest resources when that machine becomes available. This encourages
    clients to inefficiently decline resources which they could use, so that
    they will be offered better resources in the future. This may decrease
    utilization, and moreover it provides little guarantee that a valuable
    resource will be used by the framework which most values it---instead it will
    be used by whichever framework uses the least resources.
  \item If a client is waiting for a machine that is co-located with some data
    that their task needs, then Mesos encourages them to guess how likely that
    resource is to be freed up soon, and to decline new resources if this
    probability is high enough. This is actually much more efficient behavior
    than if the client behaved honestly by accepting whatever resources it could
    use, but it provides some indication that the ``guarantees'' of DRF are not
    directly applicable to the situations Mesos is intended to cover.
  \end{enumerate}

@@

@@

We will use the suggestive term “realistic conditions” to mean a situation in which any of the following occur (i.e., our negative results apply when any of these occur, and the setting is otherwise the simplest setting of static, homogeneous demands for homogeneous resources):

@@

We say that a system is “demand-based” if it elicits a demand from each user and makes assignments on the basis of the stated demands. For example, DRF, asset allocation, and CEEI all fall into this class.

@@

Suppose that there are two users and that at each time each user has enough tasks to consume all of the clusters’ resources. Tasks arrive at multiples of time T and last for T rounds. Moreover, suppose that during each interval of time T each user has an independently and uniformly randomly generated priority which is either “high” or “low.” The only characteristic of these priorities is that, all else equal, they would prefer run one extra high priority task at the expense of one fewer low priority task.

Since the scheduler is demand-based and the user’s priority levels are changing across time, the allocation must be independent of the priority of the users. So if each user receives a non-zero share of the resources, then each user must receive a non-zero share of the resources while they are in the “high” priority state and the other user is in the “low” priority state.

@@

Theorem 2. Under realistic conditions, the implementation of DRF in Mesos is not strategy-proof, Pareto efficient, envy-free, or compatible with the sharing incentive. DRF satisfies a weaker form of strategy-proofness, “look-ahead strategy proofness,” such that a user is incentivized to honestly state their demands given their beliefs about the effect of those demands on the scheduler’s future behavior.

@@

Strategy-proofness: consider the setting where a framework’s top priority is achieving some particularly valuable resource X which is currently in use, though they have other tasks which they value less. To be most extreme, suppose the user values any outcome in which they receive X higher than any outcome in which they do not. A user can arrange to receive X by strategically passing up resource offers which they could use, in order to maximize their probability of being offered the resource X when it first becomes free. DRF is only “strategy-proof” in the sense that a user who anticipates all future behavior of the scheduler will honestly state whether they want a resource in light of the consequences on the scheduler. 
Pareto-efficiency: This is a corollary of the preceding theorem. In the case of inhomogeneous resources, suppose that two users have streams of tasks, some of which run faster on resources of type A and some of which run faster on resources of type B. Then no demand-based scheduler can ensure that tasks are always placed on the resources on which they run fastest.
Envy-freeness and compatibility with sharing incentive: Suppose that one user consumes ⅓ of the resources in the cluster and another consumes ½ of the resources, but a particular 10% of the resources in the cluster are several times more valuable (to every user) than they appear. This could be based on hardware characteristics to which the scheduler is oblivious. The user who consumes ⅓ of the resources receives all resource offers first, and so claims all 10% of these resources. But this means that this user receives significantly more total value than the other user, violating both envy-freeness and the sharing incentive.

@@

Proof: %XXX, Insert proof, it’s super straightforward]
\subsection{Virtues of economic scheduling}
In this section, we show that our system possesses weaker versions of the theoretical guarantees of the theoretical properties that have been ascribed to existing schedulers. In light of the arguments in the preceding section, we hope to illustrate that the two systems are on comparable theoretical footings.

The first theorem considers the setting in which user’s willingness to pay [XXX continue]

@@

These participants generally run as daemons on the cluster.  Typically, a single
cluster would consist of up to tens of thousands of machines, each of which
running a slave process. A single master process (or multiple redundant ones
for reliability) communicates directly with all slave processes.  Each slave
process manages the tasks which run on its machine, handling requests to start
and stop various tasks from the master.

\subsection{Budgets and income}
The master keeps track of a budget for each framework: its current balance and
the amount of income it receives per second. The automatically adjusts a
framework's balance when it performs actions which require payment. Currently,
we allow a framework to incur a negative income (if it is spending beyond its
means) and also a negative balance (if negative income lasts long enough to
remove all of the savings) for simplicity.

For the master, the goal of the auction is to pick a set of tasks to run which
is feasible (i.e., does not use resource beyond what is currently available) and
which maximizes the revenue (the sum of the \emph{current price} for each of the
tasks).

@@

@@

Figures CPU-UTILIZATION and RAM-UTILIZATION show the CPU and RAM utilizations of DRF and the market-based mechanism in the high contention environment. As noted, the Google traces indicated cluster jobs were primarily CPU-bound, and hence the cluster witnesses near-maximal CPU utilization but lower RAM utilization. Figure CPU-UTILIZATION indicates that the utilization achieved by DRF and the default bidding policy are nearly identical. The average DRF utilization is approximately 98\%, while the market-based mechanism is approximately 99\%. Note the graph begins after 10 hours, when consistent maximum contention is reached. Figure RAM-UTILIZATION depicts the memory utilization, showing again the similar performance between DRF and the market-based mechanism. DRF's average memory utilization is TODO\%, while the market-based mechanism's average memory utilization is TODO\%. We do not show the disk utilization because very few jobs in the Google traces used local disk space (presumably instead relying on the Google File System).

Figure CPU-UTILIZATION-LOW shows the CPU utilizations of DRF and the market-based mechanism in a low contention environment, when no resource is fully utilized. As we can see, DRF and the market-based mechanism perform nearly identical. Both policies have an average CPU utilization of 44\%. We elide the memory and disk utilization figures because they repeat the same story, but are less interesting due to the lower RAM and disk resource requirements of jobs.

Finally, Figure JOB-COMPLETION displays the progression of workload completion under both scheduling policies. When all tasks of a job are completed, the job is considered finished. The graph shows the fraction of complete jobs as time progresses. Again, both policies result in similar behaviour, completing the workload in nearly identical fashion.

All of these results are gathered when the market-based mechanism uses a default bidding policy because the job requests provide no economic information. This shows that the market-based mechanism is ``backwards-compatible" by providing comparable performance to DRF when pricing information is not used. This result is not completely surprising though, because the default bidding policy behaves similarly to DRF. DRF awards resources to the framework with the least dominant share. Under our default bidding policy, frameworks bid a proportion of their current budget. The framework that has been awarded the least resources will have the highest budget (through less spending on resources), and will be able to bid the highest. Hence the low-resource framework will be awarded the next resource offer, similarly to DRF.

@@

Figure JOB-COMPLETION depicts the time progression of job completions, where a job is considered finished when all of its tasks are finished. As the figure depicts, the market-based mechanism is consistently ahead of DRF at all times, and finishes all jobs in about 20\% less time than DRF. Looking in detail, we find that DRF schedules only 30\% of ``fast" tasks on special resource slaves, while the market-based mechanism schedules 60\%. One might note that around the eighth hour, it appears DRF catches up to the market-based mechanism before falling behind again. This effect is caused by a long running job that the market-based mechanism schedules earlier than DRF. Its completion progresses the fraction of job completion minimally yet accounts for a long period of time. This allows DRF, which is still completing earlier short-lived tasks, to momentarily catch up in the fraction of job completion. When DRF schedules the same long-lived job at a latter point, it once again lags behind the market-based mechanism.

@@

>>>>

\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{pifont}
\newcommand{\cmark}{\ding{51}}%

@@

In section \ref{sec:design}, we describe our protocol and the most important design decisions
we made. In particular, we describe the mechanisms we used to ensure that our
protocol remains backwards-compatible, and can be applied without any change to
the clients' code.

@@

This example clearly shows that in this case, the mechanism provided by Mesos is
\emph{not} strategy-proof. Worse, it's not Pareto efficient if people are
honest, and need not be Pareto efficient if clients behave optimally. Similar
things happen in the other examples we mentioned, or in most of the similar
alternatives which we can envision:
\begin{enumerate}
  \item When one client has a constant stream of work and another has varying
    traffic, Mesos encourages both to constantly use up their allocated share of
    the network (unless they have nothing to do with the space). But an
    efficient allocation would give the client with varying workload more
    resources when it has more work. 
  \item Mesos gives a valuable resource to whatever client happens to be using
    the fewest resources when that machine becomes available. This encourages
    clients to inefficiently decline resources which they could use, so that
    they will be offered better resources in the future. This may decrease
    utilization, and moreover it provides little guarantee that a valuable
    resource will be used by the framework which most values it---instead it will
    be used by whichever framework uses the least resources.
  \item If a client is waiting for a machine that is co-located with some data
    that their task needs, then Mesos encourages them to guess how likely that
    resource is to be freed up soon, and to decline new resources if this
    probability is high enough. This is actually much more efficient behavior
    than if the client behaved honestly by accepting whatever resources it could
    use, but it provides some indication that the ``guarantees'' of DRF are not
    directly applicable to the situations Mesos is intended to cover.
  \end{enumerate}

@@

These participants generally run as daemons on the cluster.  Typically, a single
cluster would consist of up to tens of thousands of machines, each of which
runs a slave process. A single master process (or multiple redundant ones
for reliability) communicates directly with all slave processes.  Each slave
process manages the tasks which run on its machine, handling requests to start
and stop tasks from the master.

\subsection{Budgets and income}
The master keeps track of a budget for each framework: its current balance and
the amount of income it receives per second. It automatically adjusts a
framework's balance when it performs actions which require payment. For simplicity,
we allow a framework to incur a negative income and also a small temporary negative balance, although we do not accept bids from a framework with a negative balance.

For the master, the goal of the auction is to pick a set of tasks to run which
is feasible (i.e., does not use resources beyond what is currently available) and
which maximizes the revenue (the sum of the \emph{current prices} for each of the
tasks).

@@

We will use the suggestive term ``realistic conditions" to mean a situation in which any of the following occur (i.e., our negative results apply when any of these occur, and the setting is otherwise the simplest setting of static, homogeneous demands for homogeneous resources):

@@

We say that a system is ``demand-based" if it elicits a demand from each user and makes assignments on the basis of the stated demands. For example, DRF, asset allocation, and CEEI all fall into this class.

@@

Suppose that there are two users and that at each time each user has enough tasks to consume all of the clusters' resources. Tasks arrive at multiples of time T and last for T rounds. Moreover, suppose that during each interval of time T each user has an independently and uniformly randomly generated priority which is either ``high" or ``low." The only characteristic of these priorities is that, all else equal, they would prefer run one extra high priority task at the expense of one fewer low priority task.

Since the scheduler is demand-based and the user's priority levels are changing across time, the allocation must be independent of the priority of the users. So if each user receives a non-zero share of the resources, then each user must receive a non-zero share of the resources while they are in the ``high" priority state and the other user is in the ``low" priority state.

@@

Theorem 2. Under realistic conditions, the implementation of DRF in Mesos is not strategy-proof, Pareto efficient, envy-free, or compatible with the sharing incentive. DRF satisfies a weaker form of strategy-proofness, ``look-ahead strategy proofness," such that a user is incentivized to honestly state their demands given their beliefs about the effect of those demands on the scheduler's future behavior.

@@

Strategy-proofness: consider the setting where a framework's top priority is achieving some particularly valuable resource X which is currently in use, though they have other tasks which they value less. To be most extreme, suppose the user values any outcome in which they receive X higher than any outcome in which they do not. A user can arrange to receive X by strategically passing up resource offers which they could use, in order to maximize their probability of being offered the resource X when it first becomes free. DRF is only ``strategy-proof" in the sense that a user who anticipates all future behavior of the scheduler will honestly state whether they want a resource in light of the consequences on the scheduler. 
Pareto-efficiency: This is a corollary of the preceding theorem. In the case of inhomogeneous resources, suppose that two users have streams of tasks, some of which run faster on resources of type A and some of which run faster on resources of type B. Then no demand-based scheduler can ensure that tasks are always placed on the resources on which they run fastest.
Envy-freeness and compatibility with sharing incentive: Suppose that one user consumes ⅓ of the resources in the cluster and another consumes ½ of the resources, but a particular 10% of the resources in the cluster are several times more valuable (to every user) than they appear. This could be based on hardware characteristics to which the scheduler is oblivious. The user who consumes ⅓ of the resources receives all resource offers first, and so claims all 10% of these resources. But this means that this user receives significantly more total value than the other user, violating both envy-freeness and the sharing incentive.

@@

Proof: %XXX, Insert proof, it's super straightforward]
\subsection{Virtues of economic scheduling}
In this section, we show that our system possesses weaker versions of the theoretical guarantees of the theoretical properties that have been ascribed to existing schedulers. In light of the arguments in the preceding section, we hope to illustrate that the two systems are on comparable theoretical footings.

The first theorem considers the setting in which user's willingness to pay [XXX continue]

@@

\begin{table*}[t!]
\begin{center}
    \begin{tabular}{| l | l | l | l | l | l |}
    \hline
    Desired & \small DRF & \small DRF & \small Market-based  & something & something\\
    Properties & \small(simple) & \small (complex) & \small Mechanism  & something & something\\
    \hline
    \small Envy-Freeness & \cmark & \cmark  & \cmark & \cmark   & \cmark\\
    \small Pareto-Efficiency & 5 & 6  & 4  & \cmark  & \cmark\\
    \small Resource Monotonicity & 8 & 9  & 4  & \cmark  & \cmark \\
    7 & 8 & 9  & 4  & \cmark  & \cmark \\
    7 & 8 & 9  & 4  & \cmark  & \cmark\\
    7 & 8 & 9  & 4 & \cmark  & \cmark\\
    \hline
    \end{tabular}
\end{center}
  \caption{A simple table}
  \label{table:props}
\end{table*}

@@

\begin{figure}
\includegraphics[width=0.5\textwidth]{images/util_high.png}
\caption{Utilization achieved by DRF and our policy on a simulated 25-node CPU-bound cluster.  DRF's average utilization is 98\% average utilization while ours is 99\%.}
\label{img:util_high}
\end{figure}

\begin{figure}

\includegraphics[width=0.5\textwidth]{images/util_low.png}
\caption{Behavior of DRF and our policy on a simulated 100-node cluster. Both policies have average utilization of 44\%.}
\label{img:util_low}
\end{figure}

@@

Figure \ref{img:util_high} shows the CPU utilizations of DRF and the market-based mechanism in the high contention environment.  The utilizations achieved by DRF and the default bidding policy are nearly identical. The average DRF utilization is approximately 98\%, while the market-based mechanism is approximately 99\%. Note the graph begins after 10 hours, when consistent maximum contention is reached. We elide the memory utilizations graph because it is not under high contention, as the Google traces indicate jobs are CPU bound, and looks similarly to Figure \ref{img:util_low}. However, it shows again the nearly identical performance between DRF and the market-based mechanism. Similarly, we do not show the disk utilization because very few jobs in the Google traces used local disk space (presumably instead relying on a Google-internal reliable file system).

Figure \ref{img:util_low} shows the CPU utilizations of DRF and the market-based mechanism in a low contention environment, when no resource is fully utilized. As we can see, DRF and the market-based mechanism perform nearly identical. Both policies have an average CPU utilization of 44\%. Again, we omit the memory and disk utilizations figures because they repeat the same story, but are less interesting due to the lower RAM and disk resource requirements of jobs.

Finally, Figure \ref{img:reg_fairness} displays the split-wise share of cluster resources each of the five frameworks are allocated in the market-based mechanism. Under high contention, all five frameworks have queued tasks to run, and we see that each framework gets roughly a fifth of the cluster resources. This indicates the market-based mechanism allocates resources fairly , similar to DRF.

All of these results are gathered when the market-based mechanism uses a default bidding policy because the job requests provide no economic information. This shows that the market-based mechanism is ``backwards-compatible" by providing comparable utilization and fairness performances to DRF when pricing information is not used. This result is not completely surprising though, because the default bidding policy behaves similarly to DRF. DRF awards resources to the framework with the least dominant share. Under our default bidding policy, frameworks bid a proportion of their current budget. The framework that has been awarded the least resources will have the highest budget (through less spending on resources), and will be able to bid the highest. Hence the low-resource framework will be awarded the next resource offer, similarly to DRF.

@@

Figure \ref{img:inhomo} depicts the time progression of job completions, where a job is considered finished when all of its tasks are finished. As the figure depicts, the market-based mechanism is consistently ahead of DRF at all times, and finishes all jobs in about 20\% less time than DRF. Looking in detail, we find that DRF schedules only 30\% of ``fast" tasks on special resource slaves, while the market-based mechanism schedules 60\%. One might note that around the eighth hour, it appears DRF catches up to the market-based mechanism before falling behind again. This effect is caused by a long running job that the market-based mechanism schedules earlier than DRF. Its completion progresses the fraction of job completion minimally yet accounts for a long period of time. This allows DRF, which is still completing earlier short-lived tasks, to momentarily catch up in the fraction of job completion. When DRF schedules the same long-lived job at a latter point, it once again lags behind the market-based mechanism.

@@

\begin{figure}
\includegraphics[width=0.5\textwidth]{images/inhomo_times.png}
\caption{The results of simulating Mesos on a cluster of 64 machines, of which 16 have access to a special resource. DRF schedules 30\% of tasks on the optimal machine when frameworks prioritize appropriately, while a simple market-based policy schedules 60\%.}
\label{img:inhomo}
\end{figure}
